Dynamic capabilities and competitive advantage

### 2019-11-17 article4_jbr.rnw
Writing up conclusion

<<loading, results="asis", echo=FALSE >>=
setwd("/Users/larshovdanmolden/Documents/git/article4")
library(xtable)
load("./data/resart4.RData")
r <- main$resmat
rm  <- main$resmatmed
l  <- main$loadings
l <- main$loadings
#foolars=0.99
@

@@

% paper about the DC for MNC in india an dimpact on CA
%https://journals.sagepub.com/doi/abs/10.1177/0971890717701781
BLA
Extant literature on dynamic capabilities are increasingly focusing on how firm resources
are changed by dynamic capabilities as a response to changes in the environment
\citep{Schilke2018,Eriksson2014}, thus placing firm resources as a mediator between
dynamic capabilities and competetive advantage \citep{Helfat2007}. Less focus has been
placed on the nature of these underlying firm resources (ref) and how they are capturing a
rather heterogeneous bulk of routines, capabilities, assets and capital. Hence, there is a
need to disentangle the black box of firm resources to provide a more nuanced
understanding of different types of firm resources and how they interact to shape firm
competitive advantage \cite{Schilke2018}. Indeed, this gap in the literature is explicitly
emphasized by \cite{Schilke2018} where they 'see an interesting opportunity for future
work to add greater richness to our understanding of the mechanism of resource-base
change, given this mechanism’s central role in many foundational works (..) and the
diverse ways in which resource changes can potentially come about' (p. 419).

@@

@@

To test our hypotheses we then built a path-model from the latent variables we
constructed. This resulted in four models. The results from the structural model is
reported in figure \ref{fig:res} and table \ref{tab:res}. To estimate the standard errors
we ran simulations using bootstrapping and reported the simulation number, standard errors
and corresponding p-values. To evaluate the models we compared $R^2$ and $Q^2$ statistics,
as well as changes in Baysian Information Criteria (BIC) for model comparison. All these
are reported in table \ref{tab:res}. Finally, we tested the model for nonlinear effects,
endogeneity and unbserved heterogeneity using Regression Equation Specification Error Test
(RESET) , Gaussian Copula approach and Information Criteria test respectively. These
results are available in the appendix.

The estimated models is based on the following specification:

@@

\section{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The empirical models are
presented from left to right starting with the direct model. The main effect from $DC_0$
to $CA_1$, controlled for $CA_0$, firm size, age, profit margin and assets, is positive
and significant at the onset. Increasingly, by adding mediating variables as suggested by
the theory, this effect diminishes to an extent suggesting that partial or full mediation
is taking place. The effect decreases from BLA to BLA and turns insignificant. The
mediating variables $CAP_1$ and $COG_1$ are both positive and significantly related to
$CA_1$. This suggests that the effect between dynamic capabilities and competitive
advantage is mediated by these two factors separately, as well as together. These findings
both support H1 and H2. Moreover, the model becomes increasingly better both in terms if
explanatory power ($R^2$), predictive power ($Q^2$) and the Bayesian Information Criteria
(BIC).

@@

One striking finding, however, is that the improvement in the model between
model 3 and 4 is neglectable, whereas it represents a considerable improvement between 2 and 3. This
suggests that strategic cognition represents a stronger mediator than operational
capabilities. However, the difference is too small to render further analysis all but
conjecture on our part. Thus, we wanted found it necessary to explore the joint mediation
between $CAP_1$ and $COG_1$.

\subsection{Post-hoc analysis: joint mediation}

From standard production theory we know that the isoquants
between two factors of input rarely are angled (Leontief isoquants) suggesting that the
factors are total complements to eachother. Rather there are trade offs between these
factors suggesting a non-linear relationship. One way to investigate this is by means of
joint mediation where (USE MODERATED MEDIATION RATHER). 
 <<pathchild, child='./figure/medanalysisall.rnw',results="asis">>=
@

>>>>

<<loading, results="asis", echo=FALSE >>=
setwd("/Users/larshovdanmolden/Documents/git/article4")
library(xtable)
library(semPLS)
load("./data/resart4.RData")
r <- main$resmat
rm  <- main$resmatmed
l  <- main$loadings
ic <- data.frame(i[,c(1,4)]);colnames(ic) <- c("b","t")
ic$rn  <- 1:nrow(ic)
bb <- main$models$both
bm <- main$models$mmed
br <- main$models$rmed
#foolars=0.99
@

@@

% paper about the DC for MNC in india an dimpact on CA
%https://journals.sagepub.com/doi/abs/10.1177/0971890717701781

Extant literature on dynamic capabilities are increasingly focusing on how firm resources
are changed by dynamic capabilities as a response to changes in the environment
\citep{Schilke2018,Eriksson2014}, thus placing firm resources as a mediator between
dynamic capabilities and competetive advantage \citep{Helfat2007}. Less focus has been
placed on the nature of these underlying firm resources (ref) and how they are capturing a
rather heterogeneous bulk of routines, capabilities, assets and capital. Hence, there is a
need to disentangle the black box of firm resources to provide a more nuanced
understanding of different types of firm resources and how they interact to shape firm
competitive advantage \cite{Schilke2018}. Indeed, this gap in the literature is explicitly
emphasized by \cite{Schilke2018} where they 'see an interesting opportunity for future
work to add greater richness to our understanding of the mechanism of resource-base
change, given this mechanism’s central role in many foundational works (..) and the
diverse ways in which resource changes can potentially come about' (p. 419).

@@

What doe sit mean to execute DC (sense etc) and see opportunities and change underlying...

@@

To test our hypotheses we then built a path-model from the latent variables we
constructed. This resulted in four models. The results from the structural model is
reported in figure \ref{fig:res} and table \ref{tab:res}. To estimate the standard errors
we ran simulations using bootstrapping and reported the simulation number, standard errors
and corresponding p-values. To evaluate the models we compared $R^2$ and $Q^2$ statistics,
as well as changes in Baysian Information Criteria (BIC) for model comparison. All these
are reported in table \ref{tab:res}.

%% Finally, we tested the model for nonlinear effects,
%% endogeneity and unbserved heterogeneity using Regression Equation Specification Error Test
%% (RESET), Gaussian Copula approach and Information Criteria test respectively. These
%% results are available in the appendix.

The estimated models are based on the following specification:

@@

A final model includes an interaction term between $CAP_t$ and $COG_t$ to capture any
complementarity or substitution-effects these two may exhibit:

\begin{align}
  CA_{i,1} & = \alpha_0 + \alpha_1 CA_{i,0} + \omega_1 \alpha_2 CAP_{i,1}
  + \omega_2 \alpha_3 COG_{i,1} + \alpha_{4} DC_{i,0} + \alpha_5 CAP_{i,1} \times  COG_{i,1} + \mathbf{\Gamma}_m\mathbf{C}^m_{i,1} + \epsilon_i 
\end{align}

\section{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The use of PLS-SEM has
increased substantially in the realm of marketing \citep{hult2018} and strategic management
\citep{Hair2012} and is the topic of special issues in leading journals (e.g. Journal of
Business Research call for papers in 2020). In this tradition we aim to present thorough
analysis of our models and largely follows the recommendations made in recent
contributions \citep{Hair2019}.

The empirical models are presented from left to right starting with the direct model. The
main effect from $DC_0$ to $CA_1$, controlled for $CA_0$, firm size, age, profit margin
and assets, is positive and significant at the onset. Increasingly, by adding mediating
variables as suggested by the theory, this effect diminishes to an extent suggesting that
partial or full mediation is taking place. The effect decreases from 0.25 to 0.08 and
turns insignificant when adding the two mediators. This suggests a full mediation. The
mediating variables $CAP_1$ and $COG_1$ are both positive and significantly related to
$CA_1$. This suggests that the effect between dynamic capabilities and competitive
advantage is mediated by these two factors separately, as well as together. These findings
both support H1 and H2. Moreover, the model becomes increasingly better both in terms if
explanatory power ($R^2$) and predictive power ($Q^2$).

When adding the interaction term  $CAP_1 \times COG_1$ we get a significant effect. this
is compatible with the idea of behavioral ($CAP_t$) and non-behavioral ($COG_t$) objects
of change being complementary. This lends support to H3.

Our results indicate that between 37\%  and 48\% of the variation in $CA_1$ can be explained by our model,
whereas 19\% and 18\% of the variation in our mediators $CAP_1$ and $COG_t$ respectively
is explained.

@@

\subsection{Robustness and assumption validation}
The results are subject to some robustness checks not directly reported in the paper,
albeit a summary is available in the appendix. First, controlling for functional form
using a Ramsay RESET test we find support for a linear relationship between our
constructs. Second, we ran the a Variance Inflation Test (VIF) to control for
multicolinearity. None of the VIF values exceeded the suggested level indicating
multicolinearity problems. Third, we ran two separate models where we (1) included other
measures of $DC$ (using the original measure of $DC_0$), and (2) added fewer control
variables to which increased the sample size (from $N=205$ to $N=240$). Appendix 2
presents the results from all these robustness checks and assumption validations.

%% Second, running a Bayesian Copula approach, we assess the problem of
%% endogeneity in the model stemming from causes such as omitted variable bias or
%% simultaneity. The approach explicitly models the correlation between the proposed
%% endogenous variable ant the error term by means of a copula. When including a copula in
%% the estimation as a correction for endogeneity we can compare changes in the estimated
%% coefficients in terms of how they change. We tested this approach on all our three main
%% independent variables ($CAP_1$, $COG_1$ and $DC_0$) and found no support for endogeneity
%% problems in our model. However, $CAP_1$ did not fully satisfy the non-normality assumption
%% of this approach as tested using a Kolmogorov-Smirnoff test, so we cannot fully conclude
%% that endogenous effects may be at play in the model. However, the normality test did not
%% establish normality, so we believe the Bayesian Copula approach was rather efficient for
%% our purpose. Consequently, we do not believe that we have endogeneity problems in our
%% model.
%% introduced the Gaussian copula approach, which controls for endogeneity by directly
%% modeling the correlation between the endogenous variable and the error term by means of a
%% copula Model estimation in PLS-SEM draws on a three-stage approach that belongs to the
%% family of (alternating) least squares algorithms (Mateos-Aparicio 2011). Stage 1 of the
%% PLS-SEM algorithm iteratively determines indicator weights, composite scores, and path
%% coefficients. The algorithm converges when the indicator weights change only marginally
%% from one iteration to the next. Stages 2 and 3 use the final composite scores from Stage 1
%% as input for a series of ordinary least squares regressions. These regressions produce the
%% final indicator loadings, indicator weights, and path coefficients, as well as related
%% elements, such as the indirect and total effects, and R2 values (e.g., Sarstedt et
%% al. 2017a). With Simply put, the copula connects the marginal distributions of two or more
%% variables that follow any distribution (e.g., normal, non-normal). Park and Gupta (2012)
%% add a copula term to the model that represents the correlation between the endogenous
%% variable and the error term. By including this term, the effect of the endogenous
%% regressor can be estimated consistently

 <<pathchild, child='./figure/medanalysisall.rnw',results="asis">>=
@

\section{Discussion}
The aim of this paper is to theorize about how non-behavioral objects of change work as a
mechanism through which dynamic capabilities influence competitive advantage. Our results
indicate that both behavioral (\emph{operating capabilities}) and non-behavioral
(\emph{strategic cognition}) mediates the effect from dynamic capabilities on competitive
advantage. The first mechanism, dubbed 'conducting' in reference to the well established
terminology of 'resource orchestration' \citep{Sirmon2011,Helfat2007}, shows a robust and
strong function as a mediator in line with our first hypothesis. When firms exhibit strong
dynamic capabilities they are able to adapt their operating capabilites to stay
competitive. We hypothesized that this relationship would be positive and significant. Our
results indicated exactly this relationship and thus lending support to hour hypothesis.
In our example context of Microsoft Nadella did several investments in capabilities and
resources as a response to Microsofts dynamic capabilities execution. For example, he
acquired important capabilities from M\&A efforts, and upgraded the competencies regarding
cloud business and big data (Nadella). This should not be surprising given the vast
literature on indirect effects of dynamic capabilities \citep{Eriksson2014} supported by
empirical and theoretical research
\citep{Fainshmidt2016a,Protogerou2012,Karimi2015}. However, it adds to this ongoing effort
of investigating mechanisms through which dynamic capabilities work which is called for in
the literature \citep{Schilke2018}.

In this vein, our second hypothesis, and the main contribution of this paper, is much more
novel in its characteristics. Building on the important work of \cite{Verona2011} in what
they call adding a 'human side to dynamic capabilities' we suggest that non-behavioral
objects of action exists in parallel with behavioral ones. This view holds that cognition
makes a difference in creating strategic change and should thus be a part of a more
'holistic' approach to dynamic capabilities theory (ibid). The notion of cognition in
general plays an important role in determining 'the way we do things' (Vince and Gabriel
2011, ch 15). The idea of 'strategic cognition' seems to be a way of safeguarding that
organization members buy into the needed strategic change identified by execution of
dynamic capabilities. Indeed, this is also what we find in our results. Our
operationalization of 'strategic cognition' as \emph{proactiveness} makes for a proper
mediator in the focal relationship between dynamic capabilities and competitive
advantage. This is the path we have dubbed 'coaching' where dynamic capabilities rather
than orchestrating resources enables the organization to fully grasp and adapt to the proposed
strategic change. In our Microsoft example this was exactly where Nadella put the most
emphasis. By instilling a growth mindset Microsoft became a more proactive company, but
the triggering mechanism was the execution of dynamic capabilities through sensing and
seizing the opportunities they saw.

Our empirical results support the idea that \emph{operating capabilities}) and
\emph{strategic cognition} being distinctly different mechanisms through which dynamic
capabilities influence competitive advantage.

Specifically, we ask the following research question:
